Community engagement and subgroup meta-knowledge: Some factors in the soul of a community

### 2015-08-05 PaperFinalDraft.Rnw
Fixing twolevel, and putting ! in some of my is.na() expressions

In order to explore this, we study the pattern of missing data within the 2010 data set. This data set includes a large number of responses that are extremely sparse, containing almost no completed survey questions, but generally completed demographic information. For example, the question ``How likely are you to recommend this community to a friend or associate as a place to live?'' has \Sexpr{sum(is.na(sotc10$recommend)==TRUE)} missing values, and the question ``The overall quality of the colleges and universities'' has \Sexpr{sum(is.na(sotc10$college)==TRUE)} missing values. Just between those two questions, the overlap of missing values for both questions is \Sexpr{matchup1}. In other words, almost all the entries that are missing a response to the question about recommending the community are also missing a response to the question about colleges. Similarly, there are \Sexpr{matchup2} respondents missing both the question about colleges and the question ``Are you registered to vote?''

@@

# Function that takes a dataset and calculates the percentages of cases overall
findPercen <- function(dataset){
  dfall <- data.frame()
  for (i in 1:length(twolevel)){
  if(twolevel[i] %in% names(dataset)){
      tmp <- data.frame(Question=twolevel[i], prop.table(table(dataset[,twolevel[i]])))
      dfall <- rbind(dfall, tmp)
  }
  else{
    tmp <- data.frame(Question=twolevel[i], Var1 = c("No", "Yes"), Freq=c(NA, NA))
    dfall <- rbind(dfall, tmp)
  }
}
names(dfall) <- c("Question", "Response", "Freq")
return(dfall)
}
#dplyr would be much better here, but can't quite figure out how to forumulate it.

df10 <- findPercen(sotc10)
df10$Year <- 2010
df09 <- findPercen(sotc09)
df09$Year <- 2009
df08 <- findPercen(sotc08)
df08$Year <- 2008

@@

minoritiesYN09 <- sotc09 %>% 
  filter(race != " " & race != "4" & race != "Don't know" & race != "Refused") %>% 
  filter(is.na(minorities)==FALSE) %>% group_by(citystate, race != "White", minorities) %>% 
  summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2009)
names(minoritiesYN09) <- n1

minoritiesYN10 <- sotc10 %>% 
  filter(race != " " & race != "4" & race != "Don't Know" & race != "Refused") %>% 
  filter(is.na(minorities)==FALSE) %>% 
  group_by(citystate, race != "White", minorities) %>% summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2010)
names(minoritiesYN10) <- n1

@@

<<OverallSeniors>>=
seniorGroup10 <- sotc10 %>% 
  filter(is.na(seniors) == FALSE) %>% 
  group_by(citystate, age >= 62, seniors) %>% 
  summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2010)

@@

seniorGroup09 <- sotc09 %>% 
  filter(is.na(age) == FALSE) %>% 
  group_by(citystate, age >= 62, seniors) %>% 
  summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2009)
names(seniorGroup09) <- n1

@@

<<kidsbyYear>>=
kidGroup10 <- sotc10 %>% 
  filter(children == "Yes" | children == "No", is.na(families_kids) == FALSE) %>% 
  group_by(citystate, children, families_kids) %>% 
  summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2010)
names(kidGroup10) <- n1

kidGroup09 <- sotc09 %>% 
  filter(children == "Yes" | children == "No", is.na(families_kids) == FALSE) %>% 
  group_by(citystate, children, families_kids) %>% 
  summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2009)
names(kidGroup09) <- n1

@@

kidGroup08 <- sotc08 %>% 
  filter(children == "Yes" | children == "No", is.na(families_kids) == FALSE & families_kids != "(DK)" & families_kids != "(Refused)") %>% 
  group_by(citystate, children, families_kids) %>% 
  summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2008)
names(kidGroup08) <- n1

>>>>

In order to explore this, we study the pattern of missing data within the 2010 data set. This data set includes a large number of responses that are extremely sparse, containing almost no completed survey questions, but generally completed demographic information. For example, the question ``How likely are you to recommend this community to a friend or associate as a place to live?'' has \Sexpr{sum(is.na(sotc10$recommend))} missing values, and the question ``The overall quality of the colleges and universities'' has \Sexpr{sum(is.na(sotc10$college))} missing values. Just between those two questions, the overlap of missing values for both questions is \Sexpr{matchup1}. In other words, almost all the entries that are missing a response to the question about recommending the community are also missing a response to the question about colleges. Similarly, there are \Sexpr{matchup2} respondents missing both the question about colleges and the question ``Are you registered to vote?''

@@

df10 <- sotc10 %>%
  select(one_of(twolevel)) %>%
  mutate(id = 1:n()) %>%
  tidyr::gather("Question", "Response", -id) %>%
  group_by(Question, Response) %>%
  filter(!is.na(Response)) %>%
  summarise(n = n()) %>%
  mutate(Freq = n/sum(n))
df10$Year <- 2010

df09 <- sotc09 %>%
  select(one_of(twolevel[c(1:5)])) %>%
  mutate(id = 1:n()) %>%
  tidyr::gather("Question", "Response", -id) %>%
  group_by(Question, Response) %>%
  filter(!is.na(Response)) %>%
  summarise(n = n()) %>%
  mutate(Freq = n/sum(n))
df09$Year <- 2009

df08 <- sotc08 %>%
  select(one_of(twolevel[c(1:5)])) %>%
  mutate(id = 1:n()) %>%
  tidyr::gather("Question", "Response", -id) %>%
  group_by(Question, Response) %>%
  filter(!is.na(Response) & Response != "(DK)" & Response != "(Refused)") %>%
  summarise(n = n()) %>%
  mutate(Freq = n/sum(n))
df08$Year <- 2008

@@

minoritiesYN09 <- sotc09 %>% 
  filter(race != " " & race != "4" & race != "Don't know" & race != "Refused") %>% 
  filter(!is.na(minorities)) %>% group_by(citystate, race != "White", minorities) %>% 
  summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2009)
names(minoritiesYN09) <- n1

minoritiesYN10 <- sotc10 %>% 
  filter(race != " " & race != "4" & race != "Don't Know" & race != "Refused") %>% 
  filter(!is.na(minorities)) %>% 
  group_by(citystate, race != "White", minorities) %>% summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2010)
names(minoritiesYN10) <- n1

@@

<<OverallSeniors>>=
seniorGroup10 <- sotc10 %>% 
  filter(!is.na(seniors)) %>% 
  group_by(citystate, age >= 62, seniors) %>% 
  summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2010)

@@

seniorGroup09 <- sotc09 %>% 
  filter(!is.na(age)) %>% 
  group_by(citystate, age >= 62, seniors) %>% 
  summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2009)
names(seniorGroup09) <- n1

@@

<<kidsbyYear>>=
kidGroup10 <- sotc10 %>% 
  filter(children == "Yes" | children == "No", !is.na(families_kids)) %>% 
  group_by(citystate, children, families_kids) %>% 
  summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2010)
names(kidGroup10) <- n1

kidGroup09 <- sotc09 %>% 
  filter(children == "Yes" | children == "No", !is.na(families_kids)) %>% 
  group_by(citystate, children, families_kids) %>% 
  summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2009)
names(kidGroup09) <- n1

@@

kidGroup08 <- sotc08 %>% 
  filter(children == "Yes" | children == "No", !is.na(families_kids) & families_kids != "(DK)" & families_kids != "(Refused)") %>% 
  group_by(citystate, children, families_kids) %>% 
  summarise(n = n()) %>% 
  mutate(frac=prop.table(n), year=2008)
names(kidGroup08) <- n1
